### 2024-11-18 paper/3-motivation.tex
Use cref for example

% an example of Wasm control flow in official prose
\textbf{Example 1}
\begin{verbatim}
  // infinite loop
  (loop (result i32) (i32.const 42) (br 0) (unreachable) end) (unreachable)
\end{verbatim}

Example 1 is a WebAssembly code example that has \texttt{loop} and
\texttt{unreachable}
Here, result type of the \texttt{loop} is \texttt{i32} and the \texttt{loop}
has a block of three instructions: \texttt{i32.const}, \texttt{br}, and
\texttt{unreachable}.
\cref{fig:loop} is the \officialp{} of the \texttt{loop} instruction.
It says that the continuation is the start of the loop, the information is
stored in a label, and it \textit{enters} the block with the label.
The term \textit{enter} means that it pushes the label in the stack and makes
pc points to the first instruction in the block.
The \texttt{i32.const} instruction is executed first, which just pushes the
\texttt{i32} value \texttt{42} in the stack.
Then, The \texttt{br} instruction is executed.
\cref{fig:br} is the \officialp{} of the \texttt{br} instruction.
It pops the label from the stack, and makes pc points to the continuation of
the label, which is the start of the loop instruction.
As a result, the code example above is a infinite loop pushing \texttt{42}
forever without executing \texttt{unreachable} instructions.

@@

% an example of exiting label in official prose
\textbf{Example 2}
\begin{verbatim}
  // exiting label
  (loop (result i32) (i32.const 42) end) (f64.const 3.14)
\end{verbatim}

Example 2 is a WebAssembly code example that has \texttt{loop} whose block is
only \texttt{i32.const}, and \texttt{f32.const}
When the \texttt{loop} instruction is executed, it enters the block with a label.
After \texttt{i32.const} is executed, the end of the block is reached.
As a result, exiting label occurs so that the label is popped from the stack,
pc changes to the point after the end of block: \texttt{f64.const}.
As a result, after executing the code, the value 42 and 3.14 pushed to the
stack.

@@

Consider example 1 above.
If we assume some control structure $c^*$ is given, the code can be expressed
like \cref{eq:loop-1}.
\cref{fig:spectec-loop} is the \spectecp{} of the \texttt{loop} instruction.
Rather than storing the point to jump as a continuation in the label, it stores
the \texttt{loop} instruction itself in the label, and \textit{enters} the
block.
Here, \textit{enter} means that it pushes the label in the stack and executes
the instructions in the block, which means that \texttt{i32.const},
\texttt{br}, \texttt{unreachable}, \texttt{end} and \texttt{unreachable} become
the new inputs in \cref{eq:loop-2}.
When \texttt{i32.const} is excuted, it pushes 42 as \officialp{} does
in \cref{eq:loop-3}.
The value is omitted for brevity.
However, \texttt{br} behaves a bit differently.
\cref{fig:spectec-br} is the \spectecp{} of the \texttt{br} instruction.
It pops the label from the stack, removes the input instructions until the end
of the block including the \texttt{end}, considers the loop instruction in the
label as a new input instruction.
Consequently, remaining inputs are \texttt{loop}, and \texttt{unreachable}
again in \cref{eq:loop-4}.
Therefore, it explains the same behavior in a different point of view.

@@

% an example of function call in official prose
\textbf{Example 3}
\begin{verbatim}
  // function definition of $push42
  (func $push42 (result i32) (i32.const 42))

  // function call
  (call $push42) (f32.const 3.14)
\end{verbatim}

@@

Example 3 contains a function definition whose body is \texttt{i32.const}, a
\texttt{call} instruction followed by a \texttt{f32.const} instruction, which
can be expressed in \cref{eq:call-1}.
By invoking a function, it pushes a frame and enters the block with a label in
\cref{eq:call-2}.
After executing the body in \cref{eq:call-3}, an \texttt{end} is executed.
In this point, this \texttt{end} represent both the end of the block and the
function.
As a consequence, this \text{end} perform both exiting label and returning from
a function in \cref{eq:call-4}, and the instruction after the \texttt{call} is
executed in \cref{eq:call-5}
It results in the values 42 and 3.14 pushed to the stack.

@@

% an example of buggy test case
\begin{example}
\begin{verbatim}
  // function definition of $br-returning
  (func  $brAndReturning (br 0) (unreachable))

  // function call
  (call $brAndReturning)
\end{verbatim}
\label{ex:bug}
\end{example}

>>>>

% an example of Wasm control flow in official prose
\begin{example}
\label{ex:br}
\begin{verbatim}
  // infinite loop
  (loop (result i32) (i32.const 42) (br 0) (unreachable) end) (unreachable)
\end{verbatim}
\end{example}

\cref{ex:br} is a WebAssembly code example that has \texttt{loop} and
\texttt{unreachable}
Here, result type of the \texttt{loop} is \texttt{i32} and the \texttt{loop}
has a block of three instructions: \texttt{i32.const}, \texttt{br}, and
\texttt{unreachable}.
\cref{fig:loop} is the \officialp{} of the \texttt{loop} instruction.
It says that the continuation is the start of the loop, the information is
stored in a label, and it \textit{enters} the block with the label.
The term \textit{enter} means that it pushes the label in the stack and makes
pc points to the first instruction in the block.
The \texttt{i32.const} instruction is executed first, which just pushes the
\texttt{i32} value \texttt{42} in the stack.
Then, The \texttt{br} instruction is executed.
\cref{fig:br} is the \officialp{} of the \texttt{br} instruction.
It pops the label from the stack, and makes pc points to the continuation of
the label, which is the start of the loop instruction.
As a result, the code example above is a infinite loop pushing \texttt{42}
forever without executing \texttt{unreachable} instructions.

@@

% an example of exiting label in official prose
\begin{example}
\label{ex:exit}
\begin{verbatim}
  // exiting label
  (loop (result i32) (i32.const 42) end) (f64.const 3.14)
\end{verbatim}
\end{example}

\cref{ex:exit} is a WebAssembly code example that has \texttt{loop} whose block is
only \texttt{i32.const}, and \texttt{f32.const}
When the \texttt{loop} instruction is executed, it enters the block with a label.
After \texttt{i32.const} is executed, the end of the block is reached.
As a result, exiting label occurs so that the label is popped from the stack,
pc changes to the point after the end of block: \texttt{f64.const}.
As a result, after executing the code, the value 42 and 3.14 pushed to the
stack.

@@

Consider \cref{ex:br} above.
If we assume some control structure $c^*$ is given, the code can be expressed
like \cref{eq:loop-1}.
\cref{fig:spectec-loop} is the \spectecp{} of the \texttt{loop} instruction.
Rather than storing the point to jump as a continuation in the label, it stores
the \texttt{loop} instruction itself in the label, and \textit{enters} the
block.
Here, \textit{enter} means that it pushes the label in the stack and executes
the instructions in the block, which means that \texttt{i32.const},
\texttt{br}, \texttt{unreachable}, \texttt{end} and \texttt{unreachable} become
the new inputs in \cref{eq:loop-2}.
When \texttt{i32.const} is excuted, it pushes 42 as \officialp{} does
in \cref{eq:loop-3}.
The value is omitted for brevity.
However, \texttt{br} behaves a bit differently.
\cref{fig:spectec-br} is the \spectecp{} of the \texttt{br} instruction.
It pops the label from the stack, removes the input instructions until the end
of the block including the \texttt{end}, considers the loop instruction in the
label as a new input instruction.
Consequently, remaining inputs are \texttt{loop}, and \texttt{unreachable}
again in \cref{eq:loop-4}.
Therefore, it explains the same behavior in a different point of view.

@@

% an example of function call in official prose
\begin{example}
\label{ex:invoke}
\begin{verbatim}
  // function definition of $push42
  (func $push42 (result i32) (i32.const 42))

  // function call
  (call $push42) (f32.const 3.14)
\end{verbatim}
\end{example}

@@

\cref{ex:invoke} contains a function definition whose body is \texttt{i32.const}, a
\texttt{call} instruction followed by a \texttt{f32.const} instruction, which
can be expressed in \cref{eq:call-1}.
By invoking a function, it pushes a frame and enters the block with a label in
\cref{eq:call-2}.
After executing the body in \cref{eq:call-3}, an \texttt{end} is executed.
In this point, this \texttt{end} represent both the end of the block and the
function.
As a consequence, this \text{end} perform both exiting label and returning from
a function in \cref{eq:call-4}, and the instruction after the \texttt{call} is
executed in \cref{eq:call-5}
It results in the values 42 and 3.14 pushed to the stack.

@@

% an example of buggy test case
\begin{example}
\label{ex:bug}
\begin{verbatim}
  // function definition of $br-returning
  (func  $brAndReturning (br 0) (unreachable))

  // function call
  (call $brAndReturning)
\end{verbatim}
\end{example}
